---
title: "Multi-Article Classifier"
author: "Caro/Yosune"
date: "2025-05-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# 📚 Automated Classification of Multiple Articles
As part of our effort to systematically review and compare academic literature related to DMDU, we developed a framework to classify articles based on three core analytical modules:

1. Types and manifestations of uncertainty
2. Stakeholders and participatory processes
3. Implementation challenges in decision-making contexts

The purpose of this framework is to identify patterns, gaps, and emphases across the literature — for instance, whether most studies address epistemic uncertainty, whether beneficiaries are considered as stakeholders, or which kinds of implementation challenges are most commonly cited in Latin American contexts.

To make this process scalable and replicable, we constructed a classification matrix that links each category in the framework to a set of representative keywords. This matrix is editable and can be expanded to include additional fields such as methods (e.g., RDM, adaptive pathways), tools, or other relevant concepts.

## ⚙️ What Does This Multi-Article Script Do?

This script performs automated classification and standardization of multiple academic articles using a keyword-based framework. It reads and processes each article individually, matches content to pre-defined thematic categories, and generates a unified dataset for cross-article comparison.

Specifically, it:
1. Reads all .pdf, .docx, and .txt files in a target folder.
2. Extracts and cleans the text of each article.
3. Applies a classification matrix organized by module and category.
4. Counts keyword matches and calculates standardized metrics:
5. Relative weight of each category within its article and module,
6. Density of mentions per 1,000 words.
6. Produces a single, merged dataset suitable for visualization and thematic comparison across articles.
______________________________________________________

Este bloque instala y carga los paquetes necesarios para la lectura de datos, procesamiento de texto, visualización y manipulación de archivos.

This chunk installs and loads the required packages for reading data, text processing, visualization, and file handling.

``` {r Package Installation and Library Loading, echo=FALSE}
install.packages(c("readxl", "stringr", "tm", "readtext", "pdftools", "textstem", "tokenizers", "stopwords", "stringi"))

library(readxl)
library(stringr)
library(tm)
library(readtext)
library(pdftools)
library(ggplot2)
library(dplyr)
library(readr)
library(texstem)
library (tokenizers)
library(stopwords)
library(stringi)
```

Define las rutas a:
1. La matriz de clasificación en Excel, que contiene los módulos y sus palabras clave.
2. La carpeta donde se almacenan los artículos en formato .pdf, .docx o .txt.
3. Genera una lista de archivos válidos en esa carpeta.

Defines the paths to:

1. The Excel classification matrix containing modules and their keywords.
2. The folder containing the articles (.pdf, .docx, or .txt).
3. Generates a list of valid files in that folder.

``` {r Define File Paths, echo=FALSE}
matrix_path <- "data/dmdu_classification_matrix_with_keywords.xlsx"

folder_path <- "data/papers_database"

article_files <- list.files(folder_path, 
                            pattern = "\\.(pdf|docx|txt)$", 
                            full.names = TRUE, 
                            ignore.case = TRUE)
```


Este bloque:
1. Lee todas las hojas del archivo Excel y las guarda como lista de módulos.
2. Define una función para leer texto desde distintos tipos de archivo (.pdf, .docx, .txt).
3. Define una función para contar coincidencias con palabras clave.
4. Itera sobre cada artículo:
   4.1 Limpia el texto.
   4.2 Cuenta el total de palabras.
   4.3 Clasifica el contenido por módulo y categoría usando coincidencias de palabras clave.
   4.4 Calcula el porcentaje interno dentro de cada módulo.
   4.5 Guarda los resultados por artículo.


This chunk performs key tasks:
1. Reads all sheets in the Excel file and stores them as a list of modules.
2. Defines a function to read text from various file types (.pdf, .docx, .txt).
3. Defines a function to count keyword matches.
4. Iterates over each article:
   4.1 Cleans the text.
   4.2 Counts total words.
   4.3 Classifies content by module and category based on keyword matches.
   4.4 Computes the percentage within each module.
   4.5 Stores the results per article.

```{r Read the Classification Matrix, read and clean the article Text echo=FALSE}
# === 2. Read the classification matrix (each sheet is a module) ===
sheets <- excel_sheets(matrix_path)
matrix_list <- lapply(sheets, function(sheet) {
  read_excel(matrix_path, sheet = sheet)
})
names(matrix_list) <- sheets


# === 3. Function to read and clean text from files ===
get_article_text <- function(path) {
  ext <- tools::file_ext(path)
  if (ext == "txt") {
    return(tolower(readLines(path, warn = FALSE)))
  } else if (ext == "pdf") {
    return(tolower(paste(pdf_text(path), collapse = " ")))
  } else if (ext == "docx") {
    return(tolower(paste(readtext(path)$text, collapse = " ")))
  } else {
    stop("Unsupported file type.")
  }
}

# === 4. Function to count occurrences of keywords ===
match_keywords <- function(text, keywords) {
  terms <- unlist(str_split(keywords, ",\\s*"))
  sum(sapply(terms, function(term) str_count(text, fixed(trimws(term)))))
}

# === 5. Process each item in the folder ===
all_results <- list()           # Results per article
all_word_counts <- list()       #Total number of words per article

for (file in article_files) {
  
  cat("Procesando archivo:", file, "\n")
  
 # Read and clean text
  article_text <- get_article_text(file)
  article_text <- tolower(article_text)
  article_text <- removePunctuation(article_text)
  article_text <- removeWords(article_text, stopwords("en"))
  article_text <- stripWhitespace(article_text)
  
 # ===  Count number of words ===
  word_count <- str_count(article_text, "\\S+")
  article_word_counts <- data.frame(article = basename(file), total_words = word_count)
  all_word_counts[[basename(file)]] <- article_word_counts
  
  results <- list()  # Results by module
    # Process each module (sheet in Excel)
  for (sheet in names(matrix_list)) {
    df <- matrix_list[[sheet]]
    if (!"keywords" %in% colnames(df)) next  # Skip sheets without keywords
    
    module_result <- data.frame(
      category = df[[1]],
      matches = sapply(df$keywords, function(kws) match_keywords(article_text, kws)),
      stringsAsFactors = FALSE
    )
    
    # Filter categories without matches
    module_result <- module_result[module_result$matches > 0, ]
    if (nrow(module_result) == 0) next
    
    module_result$percentage <- round(100 * module_result$matches / sum(module_result$matches), 1)
    results[[sheet]] <- module_result
  }
  
  # If there are no results, create empty entry
  if (length(results) == 0) {
    combined <- data.frame(module = NA, 
                           category = NA, 
                           matches = 0, 
                           percentage = 0,
                           stringsAsFactors = FALSE)
  } else {
   # Combine results by module
    combined <- do.call(rbind, lapply(names(results), function(mod) {
      df <- results[[mod]]
      df$module <- mod
      df
    }))
  }
  combined$article <- basename(file)  # Item name
  all_results[[basename(file)]] <- combined
  
  cat("\n---", basename(file), "---\n")
  print(combined)
}

# === Consolidate total number of words per article ===
word_counts_all <- bind_rows(all_word_counts)

```

Este bloque finaliza el análisis:
1. Une todos los resultados en un solo dataframe.
2. Calcula el total de coincidencias por artículo y por módulo.
3. Une los totales y las cuentas de palabras al dataframe principal.
4. Calcula tres métricas clave para comparar entre artículos:
   4.1 prop_article: proporción del total de coincidencias que representa una categoría.
   4.2 prop_module: proporción dentro del módulo correspondiente.
   4.3 matches_per_1000w: densidad de menciones por cada 1,000 palabras.
5. Exporta un archivo CSV estandarizado listo para visualizaciones y análisis.

This block finalizes the analysis:
1. Combines all results into a single dataframe.
2. Calculates total keyword matches per article and per module.
3. Joins the totals and word counts to the main dataframe.
4. Calculates three key metrics for cross-article comparison:
   4.1 prop_article: proportion of an article's total matches.
   4.2 prop_module: proportion within the corresponding module.
   4.3 matches_per_1000w: density of mentions per 1,000 words.
5. Exports a standardized CSV file ready for visualization and analysis.

```{r Keyword Matching Function echo=TRUE }
# === 6. Consolidate results from all articles ===
final_results <- do.call(rbind, all_results)

# === STANDARDIZATION BLOCK ===
final_results$matches <- as.numeric(final_results$matches)

# 1. Total matches per item
total_matches <- final_results %>%
  group_by(article) %>%
  summarise(total_matches_article = sum(matches, na.rm = TRUE))

# 2. Total matches per module in each article
module_totals <- final_results %>%
  group_by(article, module) %>%
  summarise(module_total = sum(matches, na.rm = TRUE))

#3. Add proportions and bring it all together
final_results_std <- final_results %>%
  left_join(total_matches, by = "article") %>%
  left_join(module_totals, by = c("article", "module")) %>%
  left_join(word_counts_all, by = "article") %>%
  mutate(
    prop_article = round(100 * matches / total_matches_article, 2),
    prop_module = round(100 * matches / module_total, 2),
    matches_per_1000w = round(matches * 1000 / total_words, 2)
  )

# === Export final  CSV ===
write.csv(final_results_std, "all_article_dmdu_classification_standardized.csv", row.names = FALSE)
```

